A Data-Driven Framework for Neural Field Modeling

### 2010-12-30 ltx/first_order_framework/response_to_reviews.tex

        \item Assume that a few kernel basis functions accurately represent the actual connectivity kernel, which might not be the case when using real data;

	\emph{The examples presented in the paper use the Mexican hat connectivity kernel. The framework is also valid for other connectivity kernels provided they are homogeneous and that they can be approximated by a set of isotropic basis functions parametrized by a coefficient vector $\mathbf{\theta}$. The mathematical formulation does not restrict the shape of the kernel, where anisotropic (we assume isotropy of the kernel basis functions to make the simplification in equation~18) arbitrary shapes can be represented. Furthermore, the kernel basis functions need not be a Gaussian shape, but can be, for example, Laplacian to describe the ``wizard hat'' connectivity or B-spline functions to describe a kernel with a more compact support. For more complex basis functions some of the analytic derivations described in the Appendix may have to be performed numerically. We have added the above statements to the methods section (after equation~13) to be more explicit in describing subtleties of the method.}

        \item Do not consider the EEG/MEG lead field when constructing their mapping between the membrane voltage and the electrophysiological data (observation function)

	\emph{In response to this important point, the authors have made an effort to be clear that we are not considering scalp EEG to be a valid data source for the estimation framework. Amendments have been made to the Introduction (last paragraph), Method (after equation 9), and Discussion. We have primarily developed the framework for higher density intracranial measurements. We believe that the spatial frequency response and sampling density of extracranial EEG/MEG measurements will be insufficient to estimate the connectivity structure.}

	\emph{..., and maybe we can say the framework can be altered in the case of scalp EEG by considering the lead field in the mapping between the membrane voltage and the electrophysiological data, resulting in different observation equation (if nonlinear we need unscented transform on observation and if not the state estimation remains unchanged.) This is from PLoS review which might be useful: In much of the previous papers cited the EEG/MEG lead field is
considered. This is different of course for scalp electrodes than for intracranial electrodes but should be explicitly spelled out because this factor cannot be ignored. In fact the observation model is not quite realistic, without lead field and without amplifier characteristics.}
    \end{enumerate}

### 2010-12-30 ltx/first_order_framework/LaTeX_Manuscript.tex
added more to the response to the review and editted the manuscript

In this paper, we demonstrate for the first time how an intracortical connectivity kernel can be inferred from data, based on a variant of the \citet{Wilson1973} neural field model. This work provides a fundamental link between the theoretical advances in neural field modeling and electrophysiological data. \dean{Dean (for reviewer 1 comment 1) (this is copied from the abstract): To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters.} The paper proceeds by first describing the continuum neural field equations that are used as the cortical model. Then a finite-dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. The next section establishes conditions, using spatial frequency analysis, for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in the following section. The results for the spatial frequency analysis and parameter estimation are then presented. Finally, the implications and limitations of this framework are discussed along with planned future developments.

@@

The mapping between the membrane voltage and the electrophysiological data, denoted by $\mathbf y_t$, is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}
    \label{eq:ObservationEquation}
	y_t(\mathbf{r}_n) =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\, d\mathbf{r}'} + 
	\varepsilon_t(\mathbf{r}_n), 
\end{equation}
where $m\left(\mathbf{r}_n-\mathbf{r}'\right)$ is the observation kernel, $\mathbf{r}_n$ defines the location of the electrodes in the field, $n=1,...,n_y$ indexes the sensors and $\varepsilon_t(\mathbf{r}_n) \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes a multivariate normal distribution with mean zero and the covariance matrix $\boldsymbol{\Sigma}_{\varepsilon} = \sigma_{\varepsilon}^2\mathbf{I}$.

>>>>

In this paper, we demonstrate for the first time how an intracortical connectivity kernel can be inferred from data, based on a variant of the \citet{Wilson1973} neural field model. This work provides a fundamental link between the theoretical advances in neural field modeling and \dean{(for reviewer 1, comment 3:) high resolution intracranial} electrophysiological data. \dean{(for reviewer 1 comment 1:) (this is copied from the abstract) To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters.} The paper proceeds by first describing the continuum neural field equations that are used as the cortical model. Then a finite-dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. The next section establishes conditions, using spatial frequency analysis, for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in the following section. The results for the spatial frequency analysis and parameter estimation are then presented. Finally, the implications and limitations of this framework are discussed along with planned future developments.

@@

The mapping between the membrane voltage and the electrophysiological data, denoted by $\mathbf{y}_t$, is modelled using the observation function that incorporates sensors with a spatial extent by
\begin{equation}\label{eq:ObservationEquation}
	y_t(\mathbf{r}_n) = \int_{\Omega} { m\left(\mathbf{r}_n-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \varepsilon_t(\mathbf{r}_n), 
\end{equation}
where $m\left(\mathbf{r}_n-\mathbf{r}'\right)$ is the observation kernel, $\mathbf{r}_n$ defines the location of the electrodes in the field, $n=1,...,n_y$ indexes the sensors and $\varepsilon_t(\mathbf{r}_n) \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes a multivariate normal distribution with mean zero and the covariance matrix $\boldsymbol{\Sigma}_{\varepsilon} = \sigma_{\varepsilon}^2\mathbf{I}$. \dean{(for reviewer 1, comment 3:) Note, since we are considering intracranial measurements recorded directly from the surface or within the cortex the lead field is not modeled by the observation equation.}
